Proximal compact spaces are Corson compact

### 2014-01-08 articles/proximity-corson/proximity-corson.tex
G update to paper
Final tweaks to article

\documentclass{amsart}
\usepackage{amssymb}

@@

\title{Proximal compact spaces are Corson compact}

@@

\keywords{Proximal, absolutely proximal, Corson compact, uniform space, $W$-space}

@@

A common generalization of metric spaces is the idea of a \term{uniform space}.  A uniform space is a determined by a collection of supersets of the diagonal in the square (called ``entourages")  satisfying certain conditions.  A uniformity induces in a natural way a topology on $X$, called the \term{uniform topology}.  A topological space $X$ is \term{uniformizable} if there is  a uniformity on $X$ which generates its topology.   Any completely regular space is uniformizable.  See the next section for more complete definitions of these concepts and some of their basic properties.

Jocelyn Bell introduced the concept of proximal spaces in her doctoral dissertation while working on some uniform box product problems due to her advisor Scott Williams.  Proximal spaces are defined to be the spaces $X$ for which there is a compatible uniformity on $X$ such that the entourage picker has a winning strategy in a certain $\omega$-length game. Every metric space is easily seen to be proximal. Bell has shown that proximal spaces are collectionwise normal, countably paracompact, and have strong preservation properties, particularly, closed subpaces and  $\Sigma$-products of proximal spaces are  proximal \cite{b}.

@@

\begin{defn}
  A \term{uniform space} is a pair $\<X,\mc{D}\>$ where $X$ is a set, and $\mc{D}$ is a uniformity. A \term{uniformity} is a filter on subsets of $X^2$, called \term{entourages}, such that for each entourage $D\in \mc D$:
  \begin{itemize}
      \item $D$ is reflexive, i.e., the diagonal $\Delta=\{\<x,x\>:x\in X\}\subseteq D$.
      \item Its inverse $D^{-1}=\{\<y,x\>:\<x,y\>\in D\}\in \mc D$.
      \item There exists $\frac{1}{2}D\in\mc D$ such that
        \[
          2\left(\frac{1}{2}D\right)=\frac{1}{2}D\circ \frac{1}{2}D=\left\{\<x,z\>:\exists y\left(\<x,y\>,\<y,z\>\in \frac{1}{2}D\right)\right\}\subseteq D.
        \]
    \end{itemize}
\end{defn}

@@

\begin{defn}
  A uniform space $(X,\mathcal D)$ is said to be \term{proximal} if and only if $\pl D\win\proxgame{X}$.  A topological space is \term{proximal} iff it admits  a compatible uniformity (i.e., one which induces its topology) which is proximal.
\end{defn}

@@

Since proximal spaces are closed under closed subsets and $\Sigma$-products, it follows (as noted by Nyikos) that
every Corson compact is proximal.

@@

\begin{thebibliography}{99}
\bibitem{b}
 J. Bell,
  \emph{An infinite game with topological consequences}, preprint.
\bibitem{e}
  R. Engelking,
  \emph{General topology}, Revised and completed edition, Heldermann Verlag, Berlin, 1989.

@@

\bibitem{nproximal}
  P.J. Nyikos,
  \emph{Proximal and semi-proximal spaces}, preprint.
\end{thebibliography}

>>>>

\documentclass{amsart}
% \documentclass[preprint]{elsarticle}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}

@@

% \title{Proximal compact spaces are Corson compact\tnoteref{t1}}
% \tnotetext[t1]{2010 Mathematics Subject Classification. 54E15, 54D30, 54A20.}
\title{Proximal compact spaces are Corson compact}

% \author[aub]{S.~Clontz\fnref{fn1}}
% \ead{steven.clontz@auburn.edu}
% \author[aub]{G.~Gruenhage\fnref{fn2}}
% \ead{gruengf@auburn.edu}

% \address[aub]{Department of Mathematics, Auburn University,
%  Auburn, AL 36830}

@@

\keywords{Proximal, absolutely proximal, Corson compact, uniform space, $W$-space}
% \begin{keyword}
%   Proximal\sep absolutely proximal\sep Corson compact\sep uniform space\sep $W$-space
% \end{keyword}

@@

A common generalization of metric spaces is the idea of a \term{uniform space}.  A uniform space is a determined by a collection of supersets of the diagonal in the square (called \term{entourages})  satisfying certain conditions.  A uniformity induces in a natural way a topology on $X$, called the \term{uniform topology}.  A topological space $X$ is \term{uniformizable} if there is  a uniformity on $X$ which generates its topology.   Any completely regular space is uniformizable.  See the next section for more complete definitions of these concepts and some of their basic properties.

Jocelyn Bell introduced the concept of proximal spaces in her doctoral dissertation while working on some uniform box product problems due to her advisor Scott Williams.  Proximal spaces are defined to be the spaces $X$ for which there is a compatible uniformity on $X$ such that the entourage picker has a winning strategy in a certain $\omega$-length game. Every metric space is easily seen to be proximal. Bell has shown that proximal spaces are collectionwise normal, countably paracompact, and have strong preservation properties, particularly, closed subpaces and  $\Sigma$-products of proximal spaces are  proximal \cite{b}.  The power of these results is illustrated by the fact that the difficult result, due independently to M.E. Rudin \cite{ru} and S.P. Gul'ko \cite{gu}, that $\Sigma$-products of metrizable spaces are normal, follows as an immediate corollary.

@@

\begin{defn}
  A \term{uniform space} is a pair $\<X,\mc{D}\>$ where $X$ is a set, and $\mc{D}$ is a uniformity. A \term{uniformity} is a filter (see e.g. \cite{k}) on subsets of $X^2$, called \term{entourages}, such that for each entourage $D\in \mc D$:
  \begin{itemize}
      \item $D$ is reflexive, i.e., the diagonal $\Delta=\{\<x,x\>:x\in X\}\subseteq D$.
      \item Its inverse $D^{-1}=\{\<y,x\>:\<x,y\>\in D\}\in \mc D$.
      \item There exists $\frac{1}{2}D\in\mc D$ such that
        \[
          2\left(\frac{1}{2}D\right)=\frac{1}{2}D\circ \frac{1}{2}D=\left\{\<x,z\>:\exists y\left(\<x,y\>,\<y,z\>\in \frac{1}{2}D\right)\right\}\subseteq D.
        \]
    \end{itemize}
\end{defn}

@@

\begin{defn}
  A uniform space $\<X,\mathcal D\>$ is said to be \term{proximal} if and only if $\pl D\win\proxgame{X}$.  A topological space is \term{proximal} iff it admits  a compatible uniformity (i.e., one which induces its topology) which is proximal.
\end{defn}

@@

Since proximal spaces are closed under closed subsets and $\Sigma$-products \cite{b}, it follows (as noted by Nyikos) that
every Corson compact is proximal.

@@

\begin{thebibliography}{99}
\bibitem{b}
 J. Bell,
  \emph{An infinite game with topological consequences}, preprint.

\bibitem{e}
  R. Engelking,
  \emph{General topology}, Revised and completed edition, Heldermann Verlag, Berlin, 1989.

@@

\bibitem{gu} S.P. Gul'ko, \emph{On properties of sets lying in $\Sigma$-products}, Dokl. Akad. Nauk. SSSR 237(1977), 505--508, Soviet Math. Dokl. 18(1977), 1438--1442.

\bibitem{k}
K. Kunen,
\emph{Set theory. An introduction to independence proofs.},
Studies in Logic and the Foundations of Mathematics, 102, North-Holland Publishing Co., Amsterdam, 1983.

\bibitem{nproximal}
  P.J. Nyikos,
  \emph{Proximal and semi-proximal spaces}, preprint.

\bibitem{ru} M.E. Rudin, \emph{The shrinking property}, Canad. Math. Bull. 26(1983), no. 4, 385-388.

\end{thebibliography}
